---
title: "Solar Panel Forecast"
output:
  html_document:
    df_print: paged
---

# Project plan

What problem are we trying to address?

Predict amount of sun units needed to produce enough energy

# Setup

```{r message=FALSE, warning=FALSE, echo=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(data.table)
  library(forecast)
  library(tidyr)
  library(lubridate)
  library(dplyr) 
  library(tseries)
  library(plotly)
  library(tidyverse)
  library(nortest)
})
```

Loading the data:

```{r}
# 'sun' - sensor data, otherwise the energy produced
building_2_sun <- readRDS("data/Building 2 sun.rds")
building_2     <- readRDS("data/Building 2.rds")
building_5_sun <- readRDS("data/Building 5 sun.rds")
building_5     <- readRDS("data/Building 5.rds")
building_8_sun <- readRDS("data/Building 8 sun.rds")
building_8     <- readRDS("data/Building 8.rds")

setDT(building_2_sun)
setDT(building_2)
setDT(building_5_sun)
setDT(building_5)
setDT(building_8_sun)
setDT(building_8)

setnames(building_2_sun, "1302611", "sun")
setnames(building_2, "1490017", "energy_produced")
setnames(building_5_sun, "1328370", "sun")
setnames(building_5, "1328347", "energy_produced")
setnames(building_8_sun, "1302169", "sun")
setnames(building_8, "1498763", "energy_produced")

# weather data
weather <- fread("data/vienna.csv")
weather[,location := NULL]
weather[,date_time := as.POSIXct(date_time, format = "%d/%m/%Y")] # 31/08/2016
```

# EDA

Summary

```{r}
# Building 2
summary(building_2_sun[,sun])
summary(building_2[,energy_produced])

# Building 5
summary(building_5_sun[,sun])
summary(building_5[,energy_produced])

# Building 8
summary(building_8_sun[,sun])
summary(building_8[,energy_produced])

# Weather
summary(weather)
```

From the summary function we can see that there are no NA values. Regarding the outliers, at building 2 there are some really extreme values, but the rest seems to be fine.

## Building 2

Sun radiation:

```{r}
building_2_sun %>%
  plot_ly(
    x=~timestamp,
    y=~sun,
    type="scatter",
    mode="lines",
    line = list(color='#ff7f0e')
  )
```

Energy produced

```{r}
plot_ly(
    data=building_2,
    x=~timestamp,
    y=~energy_produced,
    type="scatter",
    mode="lines",
    line = list(color = '#1f77b4')
)
```

We can see that this data is not normal, because it has some really extreme values. So we will have to deal with outliers somehow.

## Building 5

Sun radiation:

```{r}
building_5_sun %>%
  plot_ly(
    x=~timestamp,
    y=~sun,
    type="scatter",
    mode="lines",
    line = list(color='#ff7f0e')
  )
```

Energy produced

```{r}
plot_ly(
    data=building_5,
    x=~timestamp,
    y=~energy_produced,
    type="scatter",
    mode="lines",
    line = list(color = '#1f77b4')
)
```

This looks good, and the plot for building 2 should also look similar to this.

Plotting sun radiation vs energy produced on one plot

```{r}
building_5_sun %>%
  plot_ly(
    x=~timestamp,
    y=~sun,
    type="scatter",
    mode="lines",
    name="sun radiation",
    line = list(color='#ff7f0e')
  ) %>%
  add_trace(
    inherit = F,
    data=building_5,
    x=~timestamp,
    y=~energy_produced,
    type="scatter",
    mode="lines",
    name="energy produced",
    yaxis = "y2",
    line = list(color = '#1f77b4')
  ) %>%
  layout(
    title = "Building 5",
    yaxis2 = list(
      tickfont = list(color = '#ff7f0e'), 
      overlaying = "y",
      side = "right",
      title = "second y axis - energy"
    )
  )
```

By rescaling the two data, we plotted them together. And it clearly shows that when the sun radiation is higher, there is a proportional growth in the produced energy.

## Building 8

```{r}
building_8 %>%
  plot_ly(
    x=~timestamp,
    y=~energy_produced,
    type="scatter",
    mode="lines",
    name="enery produced"
  ) %>%
  add_trace(
    inherit = F,
    data=building_8_sun,
    x=~timestamp,
    y=~sun,
    type="scatter",
    mode="lines",
    name="sun",
    yaxis = "y2"
  ) %>%
  layout(
    title = "Building 8",
    yaxis2 = list(
      tickfont = list(color = "orange"), # color of 2nd y axis
      overlaying = "y",
      side = "right",
      title = "sun radiation" # title of 2nd y axis
    )
  )
```

We can again see that the sun radiation and energy produced is tightly linked together.

Aggregated:

```{r}
building_8[, .(energy_produced=sum(energy_produced)),
              by=.(timestamp=floor_date(timestamp, "weeks"))] %>%
  plot_ly(
    x=~timestamp,
    y=~energy_produced,
    type="scatter",
    mode="lines",
    name="enery produced"
  ) %>%
  add_trace(
    inherit = F,
    data=building_8_sun[, .(sun=sum(sun)),
                        by=.(timestamp=floor_date(timestamp, "weeks"))],
    x=~timestamp,
    y=~sun,
    type="scatter",
    mode="lines",
    name="sun",
    yaxis = "y2"
  ) %>%
  layout(
    title = "Building 8",
    yaxis2 = list(
      tickfont = list(color = "orange"), # color of 2nd y axis
      overlaying = "y",
      side = "right",
      title = "sun radiation" # title of 2nd y axis
    )
  )
```

But are there maybe other factors that could influence the energy production?

## Weather

Let's try to plot how the clouds influence the produced energy:

```{r}
monthly_weather <- weather[, .(monthly_avg=mean(cloudcover)),
                           by=.(date=floor_date(date_time, "weeks"))]

monthly_weather %>%
  plot_ly(
    x=~date,
    y=~monthly_avg,
    hoverinfo= "text",
    name="Monthly cloudcover",
    type="scatter",
    mode="lines+markers",
    marker=list(size=0.5)
  ) %>%
  add_trace(
    inherit = F,
    data=building_5_sun[, .(daily_avg=mean(sun)), 
                        by=.(day=floor_date(timestamp, "weeks"))],
    x=~day,
    y=~daily_avg,
    type="scatter",
    mode="lines",
    name="Monthly average sun"
  )
```

We can see that our weather data is accurate, because when there is a higher cloudcover, there are less sunshine.

# Preprocessing

## Outliers, NAs

### Building 2

If we take a closer look at building 2:

```{r}
plot_ly(building_2, y=~energy_produced, type="box")
```

We can see that indeed there are some extreme values.

```{r}
building_2 %>%
  plot_ly(
    x=~timestamp,
    y=~energy_produced,
    type="scatter",
    mode="lines",
    name="enery produced"
  ) %>%
  add_trace(
    inherit = F,
    data=building_2_sun,
    x=~timestamp,
    y=~sun,
    type="scatter",
    mode="lines",
    opacity=0.5,
    name="sun"
  )
```

If we plot both the energy produced and the sun radiation, we can see that there is nothing extreme anomaly in the sunshine, therefore those outliers cannot be explained and most likely are wrong values.

As on the other building the range of produced energy is between 0 and 100, we have decided to cut off the outliers here based on this criteria.

```{r}
building_2 <- building_2[energy_produced < 100 & energy_produced >= 0]
```

### Weather

```{r}
colSums(is.na(weather))
```

## Merging the data

```{r}
# --------------- building 2
b2_joined <- left_join(building_2, building_2_sun, 
                       by = c("timestamp" = "timestamp"))
summary(b2_joined)
# aggregate daily

building2 <- b2_joined[, .(daily_total_sun=sum(sun), daily_total_energy=sum(energy_produced)),
                       by=.(Day=floor_date(timestamp, "days"))]
summary(building2)

building2_weather <- left_join(building2, weather, 
                               by = c("Day" = "date_time"))


# --------------- building 5
b5_joined <- left_join(building_5, building_5_sun, 
                       by = c("timestamp" = "timestamp"))
b5_joined <- na.omit(b5_joined) # remove the last day --> bc it has NAs
summary(b5_joined)
# aggregate daily
building5 <- b5_joined[, .(daily_total_sun=sum(sun), daily_total_energy=sum(energy_produced)),
                       by=.(Day=floor_date(timestamp, "days"))]
summary(building5)

building5_weather <- left_join(building5, weather, 
                               by = c("Day" = "date_time"))

# --------------- building 8
b8_joined <- left_join(building_8, building_8_sun, 
                       by = c("timestamp" = "timestamp"))
summary(b8_joined)
# aggregate daily
building8 <- b8_joined[, .(daily_total_sun=sum(sun), daily_total_energy=sum(energy_produced)),
                       by=.(Day=floor_date(timestamp, "days"))]
summary(building8)

building8_weather <- left_join(building8, weather, 
                               by = c("Day" = "date_time"))
```

# Feature engineering

Aggregating data

```{r}
############# Daily data aggregated
building_2_sun_daily <- building_2_sun[, .(daily_sum=sum(sun)),
                                         by=.(Day=floor_date(timestamp, "days"))]
building_2_daily <- building_2[, .(daily_sum=sum(energy_produced)),
                                 by=.(Day=floor_date(timestamp, "days"))]

building_5_sun_daily <- building_5_sun[, .(daily_sum=sum(sun)),
                                         by=.(Day=floor_date(timestamp, "days"))]
building_5_daily <- building_5[, .(daily_sum=sum(energy_produced)),
                                 by=.(Day=floor_date(timestamp, "days"))]

building_8_sun_daily <- building_8_sun[, .(daily_sum=sum(sun)),
                                         by=.(Day=floor_date(timestamp, "days"))]
building_8_daily <- building_8[, .(daily_sum=sum(energy_produced)),
                                by=.(Day=floor_date(timestamp, "days"))]

############# Weekly data aggregated
building_2_sun_weekly <- building_2_sun[, .(weekly_sum=sum(sun)),
                                       by=.(Week=floor_date(timestamp, "weeks"))]
building_2_weekly <- building_2[, .(weekly_sum=sum(energy_produced)),
                               by=.(Week=floor_date(timestamp, "weeks"))]

building_5_sun_weekly <- building_5_sun[, .(weekly_sum=sum(sun)),
                                       by=.(Week=floor_date(timestamp, "weeks"))]
building_5_weekly <- building_5[, .(weekly_sum=sum(energy_produced)),
                               by=.(Week=floor_date(timestamp, "weeks"))]

building_8_sun_weekly <- building_8_sun[, .(weekly_sum=sum(sun)),
                                       by=.(Week=floor_date(timestamp, "weeks"))]
building_8_weekly <- building_8[, .(weekly_sum=sum(energy_produced)),
                               by=.(Week=floor_date(timestamp, "weeks"))]

############# Monthly data aggregated
building_2_sun_monthly <- building_2_sun[, .(monthly_sum=sum(sun)),
               by=.(Month=floor_date(timestamp, "months"))]
building_2_monthly <- building_2[, .(monthly_sum=sum(energy_produced)),
               by=.(Month=floor_date(timestamp, "months"))]

building_5_sun_monthly <- building_5_sun[, .(monthly_sum=sum(sun)),
               by=.(Month=floor_date(timestamp, "months"))]
building_5_monthly <- building_5[, .(monthly_sum=sum(energy_produced)),
               by=.(Month=floor_date(timestamp, "months"))]

building_8_sun_monthly <- building_8_sun[, .(monthly_sum=sum(sun)),
              by=.(Month=floor_date(timestamp, "months"))]
building_8_monthly <- building_8[, .(monthly_sum=sum(energy_produced)),
              by=.(Month=floor_date(timestamp, "months"))]
```

# Modeling

## Time series and decomposing

```{r}
#------------------------ Time series and decomposing

####### Monthly

building_2_monthly_ts = ts(building_2_monthly[, "monthly_sum"], frequency = 12, start = c(2016,8))
autoplot(building_2_monthly_ts)
autoplot(decompose(building_2_monthly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))

building_5_monthly_ts = ts(building_5_monthly[, "monthly_sum"], frequency = 12, start = c(2016,8))
autoplot(building_5_monthly_ts)
autoplot(decompose(building_5_monthly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))

building_8_monthly_ts = ts(building_8_monthly[, "monthly_sum"], frequency = 12, start = c(2016,8))
autoplot(building_8_monthly_ts)
autoplot(decompose(building_8_monthly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))


####### Weekly

building_2_weekly_ts = ts(building_2_weekly[, "weekly_sum"], frequency = 365.25/7, start = c(2016,8))
autoplot(building_2_weekly_ts)
autoplot(decompose(building_2_weekly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))

building_5_weekly_ts = ts(building_5_weekly[, "weekly_sum"], frequency = 365.25/7, start = c(2016,8))
autoplot(building_5_weekly_ts)
autoplot(decompose(building_5_weekly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))

building_8_weekly_ts = ts(building_8_weekly[, "weekly_sum"], frequency = 365.25/7, start = c(2016,8))
autoplot(building_8_weekly_ts)
autoplot(decompose(building_8_weekly_ts, type="additive"))
```

### Checking for stationarity

```{r}
adf.test(building_2_monthly_ts)
adf.test(building_5_monthly_ts)
adf.test(building_8_monthly_ts)
```

H0: stationary

H1: not stationary

-\> all 3 are stationary

## Holt winters

```{r}
#### Monthly

# building 2
autoplot(building_2_monthly_ts)

building_2_monthly_ts.fc <- HoltWinters(building_2_monthly_ts, beta = F, gamma = F)
building_2_monthly_ts.fc$fitted

plot(building_2_monthly_ts.fc)
plot(forecast(building_2_monthly_ts.fc, h=12))

# Buiding 5
autoplot(building_5_monthly_ts)

building_5_monthly_ts.fc <- HoltWinters(building_5_monthly_ts, beta = F, gamma = F)
building_5_monthly_ts.fc$fitted

plot(building_5_monthly_ts.fc)
plot(forecast(building_5_monthly_ts.fc, h=12))

# Building 8
autoplot(building_8_monthly_ts)

building_8_monthly_ts.fc <- HoltWinters(building_8_monthly_ts, beta = F, gamma = F)
building_8_monthly_ts.fc$fitted

plot(building_8_monthly_ts.fc)
plot(forecast(building_8_monthly_ts.fc, h=12))


#### Weekly

autoplot(building_2_weekly_ts)

building_2_weekly_ts.fc <- HoltWinters(building_2_weekly_ts, beta = F, gamma = F)
building_2_weekly_ts.fc$fitted

plot(building_2_weekly_ts.fc)
plot(forecast(building_2_weekly_ts.fc, h=365.25/7))

# Buiding 5
autoplot(building_5_weekly_ts)

building_5_weekly_ts.fc <- HoltWinters(building_5_weekly_ts, beta = F, gamma = F)
building_5_weekly_ts.fc$fitted

plot(building_5_weekly_ts.fc)
plot(forecast(building_5_weekly_ts.fc, h=365.25/7))

# Building 8
autoplot(building_8_weekly_ts)

building_8_weekly_ts.fc <- HoltWinters(building_8_weekly_ts, beta = F, gamma = F)
building_8_weekly_ts.fc$fitted

plot(building_8_weekly_ts.fc)
plot(forecast(building_8_weekly_ts.fc, h=365.25/7))
```

## AR forecasting

```{r}
(building_8_weekly_ts.fc <- ar(building_8_weekly_ts))
plot(forecast(building_8_weekly_ts.fc, h=365.25/7))
```

## Holt forecasting

```{r}
(building_8_weekly_ts.fc <- holt(building_8_weekly_ts))
building_8_weekly_ts.fc$fitted

plot(building_8_weekly_ts.fc)
plot(forecast(building_8_weekly_ts.fc))
```

## Arima/sarima

```{r}
ggAcf(building_8_weekly_ts)

training <- window(building_8_weekly_ts, start=c(2016,8),end= c(2018,11))
testing <- window(building_8_weekly_ts, start=c(2018,12) )

arima_optimal <- auto.arima(training)
arima_optimal

plot(forecast(arima_optimal, h=365.25/7))

library(astsa)

sarima_forecast <- sarima.for(training, n.ahead = length(testing),
                              p=0,d=1,q=1, P=1, D=1, Q=0)
```

```{r}
arima_optimal <- auto.arima(training, max.D = 5)
arima_optimal

plot(forecast(arima_optimal, h=365.25/7))
lines(testing, col="red")
```

todo:

-   plot errors

-   auto.arima forecast with additional paramters (xreg)

    -   sun radiation

    -   sun hours

    -   cloudcover
