---
title: "Solar Panel Forecast"
output: html_notebook
---

# Project plan

What problem are we trying to address?

Predict amount of sun units needed to produce enough energy

# Setup

```{r message=FALSE, warning=FALSE, echo=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(data.table)
  library(forecast)
  library(tidyr)
  library(lubridate)
  library(dplyr) 
  library(tseries)
  library(plotly)
})
```

Loading the data:

```{r}
# 'sun' - sensor data, otherwise the energy produced
building_2_sun <- readRDS("data/Building 2 sun.rds")
building_2     <- readRDS("data/Building 2.rds")
building_5_sun <- readRDS("data/Building 5 sun.rds")
building_5     <- readRDS("data/Building 5.rds")
building_8_sun <- readRDS("data/Building 8 sun.rds")
building_8     <- readRDS("data/Building 8.rds")

setDT(building_2_sun)
setDT(building_2)
setDT(building_5_sun)
setDT(building_5)
setDT(building_8_sun)
setDT(building_8)

setnames(building_2_sun, "1302611", "sun")
setnames(building_2, "1490017", "energy_produced")
setnames(building_5_sun, "1328370", "sun")
setnames(building_5, "1328347", "energy_produced")
setnames(building_8_sun, "1302169", "sun")
setnames(building_8, "1498763", "energy_produced")
```

All the datasets here have the same structure (first column - datetime, second column - number, units of sun detected/produced)

# Preprocessing, cleansing, imputation

## Outliers, NAs

```{r}
# Building 2
summary(building_2_sun[,sun])
summary(building_2[,energy_produced])

# Building 5
summary(building_5_sun[,sun])
summary(building_5[,energy_produced])

# Building 8
summary(building_8_sun[,sun])
summary(building_8[,energy_produced])

```

From the summary function we can see that there are no NA values.
Regarding the outliers, at building 2 there are some really extreme values, but the rest seems to be fine.

If we take a closer look at building 2:

```{r}
plot_ly(building_2, y=~energy_produced, type="box")
```

We can see that indeed there are some extreme values. As on the other building the range of produced energy is between 0 and 100, we have decided to cut off the outliers here based on this criteria.

```{r}
building_2 <- building_2[energy_produced < 100 & energy_produced >= 0]
```


## Aggregating

Aggregating time series data into daily/weekly/monthly data

```{r}
############# Daily data aggregated
building_2_sun_daily <- building_2_sun[, .(daily_avg=mean(sun)),
                                         by=.(Day=floor_date(timestamp, "days"))]
building_2_daily <- building_2[, .(daily_avg=mean(energy_produced)),
                                 by=.(Day=floor_date(timestamp, "days"))]

building_5_sun_daily <- building_5_sun[, .(daily_avg=mean(sun)),
                                         by=.(Day=floor_date(timestamp, "days"))]
building_5_daily <- building_5[, .(daily_avg=mean(energy_produced)),
                                 by=.(Day=floor_date(timestamp, "days"))]

building_8_sun_daily <- building_8_sun[, .(daily_avg=mean(sun)),
                                         by=.(Day=floor_date(timestamp, "days"))]
building_8_daily <- building_8[, .(daily_avg=mean(energy_produced)),
                                by=.(Day=floor_date(timestamp, "days"))]

############# Weekly data aggregated
building_2_sun_weekly <- building_2_sun[, .(weekly_avg=mean(sun)),
                                       by=.(Week=floor_date(timestamp, "weeks"))]
building_2_weekly <- building_2[, .(weekly_avg=mean(energy_produced)),
                               by=.(Week=floor_date(timestamp, "weeks"))]

building_5_sun_weekly <- building_5_sun[, .(weekly_avg=mean(sun)),
                                       by=.(Week=floor_date(timestamp, "weeks"))]
building_5_weekly <- building_5[, .(weekly_avg=mean(energy_produced)),
                               by=.(Week=floor_date(timestamp, "weeks"))]

building_8_sun_weeks <- building_8_sun[, .(weekly_avg=mean(sun)),
                                       by=.(Week=floor_date(timestamp, "weeks"))]
building_8_weekly <- building_8[, .(weekly_avg=mean(energy_produced)),
                               by=.(Week=floor_date(timestamp, "weeks"))]

############# Monthly data aggregated
building_2_sun_monthly <- building_2_sun[, .(monthly_avg=mean(sun)),
               by=.(Month=floor_date(timestamp, "months"))]
building_2_monthly <- building_2[, .(monthly_avg=mean(energy_produced)),
               by=.(Month=floor_date(timestamp, "months"))]

building_5_sun_monthly <- building_5_sun[, .(monthly_avg=mean(sun)),
               by=.(Month=floor_date(timestamp, "months"))]
building_5_monthly <- building_5[, .(monthly_avg=mean(energy_produced)),
               by=.(Month=floor_date(timestamp, "months"))]

building_8_sun_monthly <- building_8_sun[, .(monthly_avg=mean(sun)),
              by=.(Month=floor_date(timestamp, "months"))]
building_8_monthly <- building_8[, .(monthly_avg=mean(energy_produced)),
              by=.(Month=floor_date(timestamp, "months"))]
```

## Time series and decomposing

```{r}
building_2_monthly_ts = ts(building_2_monthly[, "monthly_avg"], frequency = 12, start = c(2016,8))
autoplot(building_2_monthly_ts)
autoplot(decompose(building_2_monthly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))

building_5_monthly_ts = ts(building_5_monthly[, "monthly_avg"], frequency = 12, start = c(2016,8))
autoplot(building_5_monthly_ts)
autoplot(decompose(building_5_monthly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))

building_8_monthly_ts = ts(building_8_monthly[, "monthly_avg"], frequency = 12, start = c(2016,8))
autoplot(building_8_monthly_ts)
autoplot(decompose(building_8_monthly_ts, type="additive"))
#autoplot(decompose(energy_produced - decompose(energy_produced)$seasonal))
```


## Weather data

```{r}
weather <- read.csv("data/vienna.csv")
setDT(weather)
weather[,date_time := as.Date(date_time)]
weather[,location := NULL]
head(weather)

```

# Feature engineering
